MR-PFU: A suite of R packages for energy conversion chain analysis

### 2023-07-24 paper.Rmd
Working on paper

Historically, societal exergy analysts have
analyzed the ECCs of individual countries
using a large number of linked spreadsheets,
starting with primary and final stage data from the [IEA](https://www.iea.org/data-and-statistics/data-product/world-energy-balances)'s
world energy balances
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
Data were in varying and inconsistent formats.
The authors of this paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to data analysis was
deemed not scalable.
We created a suite of `R` packages to create a database 
of PFU data (the PFU Database)
for all countries in the world from 1960--2020.
The packages enable, for the first time, scalable SEA.
This paper describes the design of those packages and
demonstrates briefly their use.

@@

Further development followed from selection of the PSUT data format.
First, the matrix-based PSUT framework brought the challenge that 
different countries and years 
have varying energy energy carriers (products) and
varying energy conversion machines (industries),
meaning that PSUT matrices for different countries and years 
have differing row and column names.
To get around this challenge, we created the
([matsbyname](https://github.com/MatthewHeun/matsbyname/)) 
package which enables matrix mathematics 
that respects matrix row and column names, 
inserting rows or columns of `0`s when needed.
Second,
we knew it would be convenient to perform _matrix_ mathematics 
as easily as _scalar_ mathematics
in `R` data frames
using the [tidyverse](https://www.tidyverse.org) syntax. 
We developed the [matsindf](https://matthewheun.github.io/matsindf/) package 
to enable this functionality.
Finally, manipulating row and column names proved to be a challenge, 
especially for matrices in PSUT data frames, so 
we developed the [RCLabels](https://matthewheun.github.io/RCLabels/) package
for that purpose.
The table below summarizes these packages, 
all of which are generally useful and available on 
[CRAN](https://cran.r-project.org).

>>>>

Historically, societal exergy analysts have
analyzed the ECCs of individual countries
using a large number of linked spreadsheets,
starting with primary and final stage data from the [IEA](https://www.iea.org/data-and-statistics/data-product/world-energy-balances)'s
world energy balances
[@Ayres:2003ec; @Serrenho:2014aa; @Brockway:2014aa; @Brockway:2015aa]. 
Data were in varying and inconsistent formats.
The authors of this paper and others in the field
wanted to expand SEA to cover all countries, but
the spreadsheet approach to data analysis was
deemed not scalable.
We created a suite of `R` packages to create a database^[Strictly speaking, we create data frames of matrices, 
not a structured database.]
of PFU data (the PFU Database)
for all countries in the world from 1960--2020.
The packages enable, for the first time, scalable SEA.
This paper describes the design of those packages and
demonstrates briefly their use.

@@

Further development followed from selection of the PSUT data format.
First, the matrix-based PSUT framework brought the challenge that 
different countries and years 
have varying energy energy carriers (products) and
varying energy conversion machines (industries),
meaning that PSUT matrices for different countries and years 
have differing sizes and differing row and column names.
To get around this challenge, we created the
([matsbyname](https://github.com/MatthewHeun/matsbyname/)) 
package which enables matrix mathematics 
that respects matrix row and column names, 
inserting rows or columns of `0`s when needed.
Second,
we knew it would be convenient to perform _matrix_ mathematics 
as easily as _scalar_ mathematics
in `R` data frames
using the [tidyverse](https://www.tidyverse.org) syntax. 
We developed the [matsindf](https://matthewheun.github.io/matsindf/) package 
to enable this functionality.
Finally, manipulating row and column names proved to be a challenge, 
especially for matrices in PSUT data frames, so 
we developed the [RCLabels](https://matthewheun.github.io/RCLabels/) package
for that purpose.
The table below summarizes these packages, 
all of which are generally useful and available on 
[CRAN](https://cran.r-project.org).
